Modeling Two-Dimensional Vesicle Hydrodynamics with Hydrophobic Attraction Potential Simulations

### 2021-08-13 doc/main_jfm.tex

\documentclass[lineno]{jfm}

@@

@@

@@

\end{document}

>>>>

<<<<<<< HEAD
\documentclass[lineno]{jfm}

\section{\label{intro}Introduction}
Described by de Gennes as ``another animal in soft matter physics", the Janus particle (often a spherical particle with a hydrophobic and a hydrophilic hemisphere) exhibits complex aggregate, clustering and self-assembly into meso- and macroscopic structures that are relevant to a wide range of applications in biology and bioengineering. Whether it is surface chemistry or polarity under an external field, the dynamics of Janus particles is inevitably the combination of long-range hydrodynamics interactions in the presence of short- and intermediate-range particle-particle interactions.

Recently Fu {\it et al.} illustrated that the hydrophobic interactions between Janus particles in a viscous solvent can be used as a coarse-grained model to capture the mechanics of an elastic bilayer membrane of amphiphilic macromolecules such as lipids \cite{Fu2020_SIAM}. Depending on the total number of Janus particles and their geometry, in a viscous solvent, Janus particles are found to aggregate to form a micelle, a patch of bilayer membrane with open ends, and a self-enclosed bilayer membrane (a vesicle). Using a hybrid continuum model for the interactions between amphiphilic particles in a viscous solvent and formulating it in a boundary integral formulation, Fu {\it et al.} showed that the membrane granularity during fusion and fission of bilayers can be accurately captured. 
%This boundary integral formulation has inspired more recent work

In molecular dynamics (MD) simulations, the lipid bilayer membranes are often approximated by an implicit solvent coarse-grained model for lipids (as in the Martini code), where their hydrodynamic interactions are simplified by an implicit solvent. The long-range hydrodynamic interactions and the induced long-range lipid interactions can be captured by combining the Dry Martini with a lattice Boltzmann MD \cite{Brandner2019_SciRep}. Brandner {\it et al.} used this methodology to simulate the hydrodynamics of a nano-sized vesicle under a shear flow, and reproduced the well-know vesicle hydrodynamics  such as tank-treading, tumbling and inter-leaflet slippage (reported from simulations of the Helfrich-type continuum model).

In this work we extend the hybrid continuum model in \cite{Fu2020_SIAM} for the Janus suspension to incorporate the collective hydrodynamics of a Janus suspension under various flows: a linear shear flow, an elongation flow, and a Poiseuille flow. We demonstrate that a vesicle made of Janus particles does behave like a vesicle in these flowing conditions. Furthermore we also show that the hybrid continuum model allows us to investigate the inter-leaflet slippage and rupture of a bilayer membrane due to an imposed flow.

This paper is organized as follows. In \S~\ref{sec:governing_eqs} we present the formulation for the Janus particles in a viscous fluid in the zero-Reynolds number regime. In \S~\ref{sec:IEM} we extend the hybrid continuum model for a Janus suspension to include the effects of a far-field flow via the mobility problem formulation. In \S~\ref{validation} we present validation of our model, and in \S~\ref{results} we present simulation results for a vesicle of Janus bilayer membrane under various flowing conditions. Finally we provide discussion and outlook for future directions in \S~\ref{conclusion}.

\section{Governing  Equations\label{sec:governing_eqs}}
\subsection{\label{mobility}Mobility Problem}
We consider a suspension of $N_b$ particles suspended in a
two-dimensional unbounded domain. The boundary of each particle is
denoted by $\Gamma_i$ so that $\bd \Omega = \Gamma_1 \cup \Gamma_2 \cup
\cdots \cup \Gamma_{N_b}$. Assuming the flow is Stokesian, the governing
equations are
\begin{alignat}{3}
  -\mu \Delta \uu + \nabla p &= \mathbf{0}, 
    && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
%
where $\uu$ is the velocity, $\mu$ is viscosity, $p$ is the pressure,
and $\uu_\infty$ is the background flow. Since each particle $\Gamma_i$
with center $\aa_i$ is a rigid body, its velocity satisfies 
\begin{align}
  \vv(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad 
    \xx \in \Gamma_i,
\end{align}
where $\vv_i$ is its translational velocity and $\omega_i$ is its
angular velocity. Here, $\langle x, y \rangle^{\perp} = \langle -y, x
\rangle$. Therefore, the no-slip boundary condition on each particle is
\begin{align}
  \uu(\xx) = \vv_i + \omega_i (\xx - \aa_i)^\perp, \quad
    \xx \in \Gamma_i.
\end{align}
To determine the translational and angular velocities of each particle,
we assume some imposed forces $\FF_i$ and torques $T_i$ acting on each
particle. Since the small particles are intertialess, force balance
gives 
\begin{alignat}{2}
  \label{eqn:force}
  \FF_i &+ \int_{\Gamma_i} \ssigma \cdot \nnu_i \, ds = \mathbf{0},
  && i=1,\ldots,N_b,\\
  \label{eqn:torque}
  T_i &+ \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
    (\ssigma \cdot \nnu_i) \, ds = 0, \qquad && i=1,\ldots,N_b,
\end{alignat}
where $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor). The
process of finding the translational and angular velocities given a
force and torque is referred to as the mobility problem.

In order to examine the dynamics of the Janus-vesicle using the HAP coupling with hydrodynamics interactions, for the target bilayer, a circular initial shape is chosen to perform the relaxation run. 
This vesicle-like structure consists of $N=58$ disks with particle radius $r=0.5$ and the starting diameter of 
the vesicle is about $8.5$ length units.
The starting center of mass position of the Janus-vesicle is pinned at the origin and particle velocities are tracked to determine the initial configuration for later use such as single and multiple vesicle simulations. 
This will prevent some numerical instability issues when the minimum of the total energy is not achieved. 
For instance, the dynamics of the bilayer under a shear flow may reach different outcomes.
Figure~\ref{figure2}(a) shows that the magnitudes of mean velocity components ${\bf v}=\{v_x,v_y\}$ decrease exponentially with an approximation decaying rate $\sim0.005$. In the same figure, the variation of the mean particle orientation $\theta$ also decreases over a period of time as expected from the self-assembly property of the model. This implies that a local energy minimum and the equilibrium configuration are 
nearly attained. 
The time step size $\Delta t=0.2$ is used for all simulations in this work including the relaxation run.
Figure~\ref{figure2}(b) shows the configuration difference between starting and ending frames in the relaxation run. The Figure~\ref{figure2}(c)-(d) demonstrate the fluid pressure profile of two configurations. The pressure inside the permeable Janus-vesicle decreases as expected.

Following the relaxation experiments of the Janus-vesicle, this section will focus on the behavior of vesicles 
under a number of flow types. To extract physical quantities of the target structure using the proposed model, the scaling laws and the dimensionless units are given as follows.
The length unit is scaled by the length of phospholipids $l=2.5$ nm and the force is scaled from the 
interfacial tension $\gamma=4.1$ pN nm$^{-1}$ from ~\cite{Ryham16}. We then be able to calculate the corresponding 
time scale $\tau = 0.38$ ns. With the same framework, \cite{Fu20} provides the bending rigidity 
$\kappa=8.51$ kT.

For the $N$-body Janus-vesicle, we denote a constant initial radius $R_0=\sqrt{A_0/4\pi}$ with initial area $A_0$ and the dimensionless shear rate $\chi$ is referred from~\cite{Finken08} and~\cite{Shaqfeh11},
\begin{align}
  \chi = \dot\gamma \frac{\mu R_0^2}{\kappa},
\end{align}
%
where $\dot\gamma$ the dimensional shear rate, and $\mu$ the fluid viscosity.

To study the deformation of a closed bilayer structure under a shear flow using the proposed algorithm, 
we adopt a widely tested numerical simulation in both continuum and molecular dynamics models, single vesicle under a shear flow. The well known results indicate that the vesicle does tank-treading when there is no viscosity contrast in both inside and outside of the structure.
Given a set of various shear rates, with no viscosity difference between internal and external fluids, 
the vesicle-like granular structure is initially placed in a quiescent flow for obtaining an equilibrium shape.

The snapshots of the simulation for single vesicle in a shear flow are shown in Figure~\ref{figure3}.
Panel (a) shows the initial circular shape obtained from the relaxation run and the deformed 
configuration of the vesicle after a period of time ($t=4000$) is in panel (c). The color presented in the figure 
is the strength of the action due to the hydrophobic attraction. From weak to strong action field, the color
varies from blue to red.
To demonstrate the tank-treading behavior of the vesicle, panels (b) and (d) provide the streamlines of the 
fluid velocity over the whole computational domain when the shear flow passes through the structure. The tank-treading motion of the structure is then observed, for instance, the particle velocities in panel (d) indicate a clockwise tank-treading movement.

\begin{figure}
\centering
\includegraphics[width=0.3\textwidth]{N58_0.pdf}
\includegraphics[width=0.3\textwidth]{N58_vel_0.pdf}\\
\includegraphics[width=0.3\textwidth]{N58_20000.pdf}
\includegraphics[width=0.3\textwidth]{N58_vel_20000.pdf}
  \caption{Vesicle under a shear flow where the shear rate $\dot\gamma=0.0025$: 
  (a) Initial configuration of 58-body vesicle; 
  (b) The fluid velocity plot with respect to the configuration in (a); Streamlines are plotted in the background
  as solid curves. Red arrows represent the particle velocities.
  (c) Deformed vesicle after $20000$ steps. The tank-treading motion occurs after reaching the state shown in panel (c).
  (d) The fluid velocity plot with respect to the configuration in (c);
  The background colors in panels (a) and (c) represent the strength of the hydrophobic attraction potential. The blue region shows no action and the red region with the bilayer gives the strongest action.
  }
    \label{figure3}
\end{figure}
%

In order to extract more physical properties of the proposed model, we track the total length of the bilayer structure, the reduced area over the time, and the excess length. Here the enclosed area and the length of the structure are calculated from the midplane. 
These results are shown in Figure~\ref{figure4} where we compare results from a set of shear rates $\dot\gamma=\{0.002,0.0025,0.003,0.0035\}$. 
The reduced area is given by $A^* = 4\pi A/L^2$ where $A$ is the vesicle area and $L$ is the arc length of the vesicle with respect to the midplane. 
Furthermore, the excess length is given by $\Delta=L/\sqrt{A^*/\pi}-2\pi$.
Since the starting configuration is close to a circular shape, the reduced area decreases from a value very close to $1$. It is clear that a slower shear rate generates a higher reduced area and has a faster convergence. 
With a converging total arc length, a slower shear rate also produces a smaller excess length.
The small oscillations in all results are due to the granularity of the bilayer which tend to having a slightly non-smooth curve in simulations. Overall the total lengths of the bilayer have the same converging value in all test shear rates. Moreover, based on the 
numerical experiments, we observe that the reduce area converges relatively early when the shear rate is lower than $\dot\gamma=0.003$.

An interesting phenomenon also rises up in the proposed setup where the inter-monolayer slip occurs 
between two leaflets. When a shear rate is large enough, take $\dot\gamma = 0.005$ as an example, 
Figure~\ref{figure5} shows the schematic of the test. A particle pair is initially marked and we then track the positions and mean angle $\theta(t)$ between particle relative position with respect to the vesicle center over the time. It is clear to see that the outer leaflet always has larger 
tangential velocity than the one from inner leaflet and the mean angle $\theta(t)$ increases. 
This inter-leaflet sliding phenomenon is consistent with what has been seen
in articles in molecular dynamics simulations. %%% need some citations

\begin{figure}
\begin{center}
\hspace{-0.6cm}
\includegraphics[height=2in]{ReducedArea.pdf}
\hspace{0.6cm}
\includegraphics[height=2in]{DecayRate.eps}\\
\includegraphics[height=2in]{ArcLength.pdf}
\includegraphics[height=2in]{ExcLength.pdf}
\end{center} 
  \caption{(a) Total arc lengths of the midplanes over time in different shear rates; (b) From exponential fits in panel (a), we obtain the decay rate for a number of different shear rates;
  (c) Reduced area transitions over time. (d) Excess length of the midplanes.
A set of shear rates $\dot\gamma=\{0.002,0.0025,0.003,0.0035\}$ is used in simulations. 
A slower shear rate generates a faster convergence in the reduced area and smaller excess length. 
  }
    \label{figure4}
\end{figure}

\todo[inline]{Figure/Table of the friction coefficient for different shear rates (needs to revise)}

The intermonolayer friction coefficient of the Janus-vesicle is calculated from the ratio of friction force and the 
slip velocity. This quantity is given by 
\begin{equation}
b = \frac{F_f}{v^\parallel} = \frac{F_f}{\int_{\Sigma_M} R \dot\theta dS}.
\end{equation}
%
where $v^\parallel$ represents the tangential velocity at the midplane and $\Sigma_M$ the curve determined by the midplane.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\textwidth]{Slip.eps}
\end{center} 
  \caption{The schematic of the intermonolayer slip over a period of time. Taking shear rate $\dot\gamma=0.005$ as an example, we denote the mean angle between each particle pair $\theta(t)$ where the angle is calculated from the relative position of particles using the vesicle center of mass position. The configuration plots have two fixed particles marked in yellow and blue and labeled with the corresponding time $t=\{2000,5200,9000\}$. The midplane curve is denoted as $\Sigma_M$.
  }
    \label{figure5}
\end{figure}

\begin{table}
\caption{Friction Coefficients}
\centering    
\begin{tabular}{c c c c c c c c }
\hline                    
$\dot\gamma$ & 0.0020 & 0.0025 & 0.0030 & 0.0035 & 0.0040 & 0.0045 & 0.0050\\
\hline 
$F_{f}$ &  &  &  &  &  &  &   \\[1ex]
$\dot\theta$             &  &  &  &  &  &  & \\ [1ex]
$b$             &  &  &  &  &  &  & \\ 
\hline    
\end{tabular} 
\label{table1}
\end{table}

The background flow for this set of simulations is given by

\begin{equation}
{\bf v}_\infty = v_{max}\left[ 1 - \left( \frac{y}{wR_0}\right)^2 \right]{\bf e}_x,
\end{equation}
%
where $v_{max}$ is the flow strength and $w$ controls the shape of the flow. $R_0$ is defined as the radius of the initial vesicle midplane. Figure~\ref{figure6} gives a final configuration for one specific case where the circular vesicle is initially placed above the $x$-axis.

\begin{figure}
\centering
\includegraphics[height=2in]{parabolic.eps}
  \caption{Vesicle under a parabolic flow where the strength of the flow $v_{max}=8.0$ and the control parameter $w=10$. The configuration is the state when $t=10000$.
  }
    \label{figure6}
\end{figure}

With the choice of a large shear rate, a pore formation or a complete membrane rupture can be observed during simulations. Figure~\ref{figure7} shows a demonstration of a vesicle under a shear flow with shear rate $\dot\gamma=0.1$. An interesting phenomenon occurs in panels (c) and (d) where two separated bilayers move along the background shear flow symmetrically. Consequently, a periodic motion for bilayers may occur 
with the choices of shear rates. One can expect that these two layers will flow far apart from each other for much higher shear rates.

\begin{figure}
\centering
\includegraphics[height=2in]{N58_rupt_0.pdf}
\includegraphics[height=2in]{N58_rupt_200.pdf}
\\
\includegraphics[height=2in]{N58_rupt_400.pdf}
\includegraphics[height=2in]{N58_rupt_600.pdf}
  \caption{Membrane rupture occurs when the shear rate is beyond a certain threshold. Here the shear rate is $\dot\gamma = 0.1$. Starting with a circular shape (panel (a)), the vesicle is stretched by the background flow and multiple pores appear (panel (b)). (c)-(d) are the later states during the simulation.
  }
    \label{figure7}
\end{figure}

\begin{figure}
\centering
\includegraphics[height=2in]{N116_shear_0.pdf}
\includegraphics[height=2in]{N116_shear_2500.pdf}\\
\includegraphics[height=2in]{N116_shear_5000.pdf}
\includegraphics[height=2in]{N116_shear_7500.pdf}
  \caption{Two vesicles under a shear flow where the dimensioneless shear rate is $\dot\gamma=0.005$; The initial configuration is shown in panel (a). The background color indicates the magnitude of fluid pressure. From dark red to dark blue, it represents the the pressure varies from high to low values. The color bar is valid for all panels. Streamlines are plotted in purple. The insets of all panels are generated from the results of continuum modeling simulations and the streamlines have a perfect agreement to the numerical results.
  }
    \label{figure9}
\end{figure}

\subsubsection{Shear Flow}
Two vesicles in a shear flow is presented in Figures~\ref{figure9} where centers of mass are placed on the same horizontal level. With the same starting state as one in the extensional flow, panels (b)-(d) in Figure~\ref{figure8} are the observed phenomena when two vesicles are in a shear flow where the flow is given by $\dot x = \dot\gamma y$. 
An interesting event occurs in the simulation that a rotational behavior with an adhesive effect can be observed and tuned. In other words, the adhesive behavior will be absent when two vesicles are centered 
at a different level and are well separated in $y$-direction.

\begin{figure}
\centering
\includegraphics[height=2in]{N116_ext_0.pdf}
\includegraphics[height=2in]{N116_ext_2000.pdf}\\
\includegraphics[height=2in]{N116_ext_4000.pdf}
\includegraphics[height=2in]{N116_ext_6500.pdf}
  \caption{Two vesicles under an extensional flow where the dimensioneless shear rate is $\dot\gamma=0.005$; The initial configuration is shown in panel (a). The background color indicates the magnitude of fluid pressure. From dark red to dark blue, it represents the the pressure varies from high to low values. The color bar is valid for all panels. Streamlines are plotted in purple. The insets show the results produced from using continuum model that the streamlines have perfect agreements with the present results.
  }
    \label{figure8}
\end{figure}

Figures~\ref{figure8} show the numerical results of simulations when two vesicles are in an extensional flow. Starting from the initial configuration (shown in panel (a)), panels (b)-(d) show the transitions of two vesicles moving under an extensional flow. The extensional flow is stretching along the $y$ direction and squeezing in $x$ direction with shear rate $\dot\gamma=0.005$. Since two bilayers are close enough
from the beginning of the run, two vesicles are first pushed toward each other then separate along the streamline. The short range repulsion plays an important role to avoid structure collisions and 
this result can be compared with some findings in continuum modeling simulations.
The pressure plots in Figure~\ref{figure8} provide a realistic result where the pressure is much higher in the gap between two nearby vesicles.

%\section{\label{A2} Comparison between Continuum Model and the Proposed Model}
%
%\begin{figure}
%\centering
%\includegraphics[height=1.2in]{N116_cont_a.pdf}
%\includegraphics[height=1.2in]{N116_cont_b.pdf}
%\includegraphics[height=1.2in]{N116_cont_c.pdf}
%\includegraphics[height=1.2in]{N116_cont_d.pdf}
%  \caption{Two vesicles under an extensional flow where the dimensioneless shear rate is $\dot\gamma=0.005$; Continuum modeling.....
%  }
%    \label{figureB}
%\end{figure}

\end{document}
=======
\documentclass[lineno]{jfm}

@@

%\todo[inline]{Introduce dimensionless shear rate here (and check the numbers)}

@@

%\todo[inline]{Make figure of the paths of the two centroids....}

@@

\end{document}
>>>>>>> 792ef765a54d5589595223dd98b804c25f39cde8
